---
title: "UK Online Retail"
author: "pravin kumar"
date: "15/01/2022"
output: html_document
---

A UK based retail client wants to look into their transactional data sets and perform analytics to increase sales and improve customer loyalty. You are provided a sample data (attached with this mail) for your peruse. You are requested to look into this data set and use various analytical techniques/ algorithms to understand it better and come up with insights that business can use. 

```{r required packages}
library(readxl)
library(plyr)
library(dplyr)
library(lubridate)
library(ggplot2)
```

```{r importing data}
mydata <- read_xlsx("Online Retail.xlsx")
mydata <- data.frame(mydata)
data <- mydata
str(data)
summary(data)
```

There are entries which are not transactional in nature. We should be removing those non-transactional entries to focus more on the analysis.

```{r}
data2 <- data[which(data$UnitPrice > 0),]
data2 <- subset(data2, StockCode!="D" & StockCode!="DOT" & StockCode!="S" & StockCode!="POST" & StockCode!="M" & StockCode!="C2" & StockCode!="AMAZONFEE" & StockCode!="B" & StockCode!="BANK CHARGES" & StockCode!="CRUK" & StockCode!="m" & StockCode!="PADS")
giftcardrows <- as.vector(subset(data2, subset = grepl(glob2rx("gift*"), StockCode))[,2])
data2 <- data2[-which(data2$StockCode %in% giftcardrows),]
data2[,'Description'] <- gsub(",","",data2[,'Description'])
```

We are adding new column called Canceled which will contain boolean values for canceled items. If order is canceled, the value is Yes and if order is successful, then value is No.

```{r}
data2$Canceled <- ifelse(data2$Quantity<0,"Yes", "No" )

```

We are using RFM technique to segment customers. Since RFM uses customer ID to put customer in different segments, we might want to remove rows which do not have customer ID mentioned.

```{r}
rfmdata <- data2
colSums(is.na(rfmdata))
rfmdata <-na.omit(rfmdata)
```
Adding Total column

```{r}
rfmdata$Total <- rfmdata$Quantity * rfmdata$UnitPrice
rfmdata$InvoiceDate <- as.Date(rfmdata$InvoiceDate, "%Y-%m-%d")
head(rfmdata)

summary(rfmdata)
```

RFM Analysis
```{r}
analysis_date <- max(rfmdata$InvoiceDate)
rfm_df <- rfmdata %>% group_by(CustomerID) %>% dplyr::summarise(Recency = as.numeric(analysis_date- max(InvoiceDate)), Frequency = n(), Monetary = sum(Total))
nrow(rfm_df) #total number of distinct customerID
```

Plotting Recency, Frequency and Monetary
```{r}
r <- ggplot(rfm_df) +geom_density(aes(x= Recency))
f <- ggplot(rfm_df) +geom_density(aes(x = Frequency))
m <- ggplot(rfm_df) +geom_density(aes(x = Monetary))
grid.arrange(r, f, m, nrow = 3)
```
Putting classifiers to segment customers
```{r}
rfm_df$R_score <- 0
rfm_df$R_score[rfm_df$Recency >= 141.50] <- 1
rfm_df$R_score[rfm_df$Recency >= 50.00 & rfm_df$Recency <141.50] <- 2
rfm_df$R_score[rfm_df$Recency >= 17.00 & rfm_df$Recency <50.00] <- 3
rfm_df$R_score[rfm_df$Recency < 17.00] <- 4
rfm_df$F_score<- 0
rfm_df$F_score[rfm_df$Frequency >=100] <- 4
rfm_df$F_score[rfm_df$Frequency <100 & rfm_df$Frequency >= 41] <- 3
rfm_df$F_score[rfm_df$Frequency <41 & rfm_df$Frequency >= 17] <- 2
rfm_df$F_score[rfm_df$Frequency <17] <- 1
rfm_df$M_score <- 0
rfm_df$M_score[rfm_df$Monetary >= 1661] <- 4
rfm_df$M_score[rfm_df$Monetary < 1661 & rfm_df$Monetary >= 674.5] <- 3
rfm_df$M_score[rfm_df$Monetary >= 307.2 & rfm_df$Monetary < 674.5] <- 2
rfm_df$M_score[rfm_df$Monetary <307.2] <- 1

rfm_df <- rfm_df %>% mutate(RFM_score = 100 *R_score +10 * F_score + M_score)

rfm_df$Segment <- "0"
rfm_df$Segment[which(rfm_df$RFM_score %in% c(444,434,443, 344, 442, 244, 424, 441  ))] <-"Loyalists"
rfm_df$Segment[which(rfm_df$RFM_score %in% c(332,333,342, 343, 334, 412,413,414,431,432,441,421,422,423, 424, 433 ))] <- "Potential Loyalists"
rfm_df$Segment[which(rfm_df$RFM_score %in% c(233,234, 241,311, 312, 313,314,321,322,323,324, 331,  341))] <- "Promising"
rfm_df$Segment[which(rfm_df$RFM_score %in% c(124, 133, 134, 142, 143, 144, 214,224,234, 242, 243, 232 ))] <- "Hesitant"
rfm_df$Segment[which(rfm_df$RFM_score %in% c(122, 123,131 ,132, 141, 212, 213, 221, 222, 223, 231 ))] <- "Need attention"
rfm_df$Segment[which(rfm_df$RFM_score %in% c(111, 112, 113, 114, 121, 131, 211, 311, 411 ))] <-"Detractors"
```
 
 Plotting segments
```{r}
table(rfm_df$Segment)
ggplot(rfm_df) + geom_bar(aes(x = Segment, fill = Segment))+theme(axis.text.x=element_text(angle=90,hjust=1)) +labs(title = "Barplot for Segments of customers")
```

Based on above described segments, company can prepare strategy to convert promising, potential loyalists into Loyalists by offering them discounts, gift cards to maintain the positive relationship. 
For customers in need attention and hesitant segment, company can start campaign to focus on these segments by listening to their feedback and working more on what they has to say. It should be taken into consideration that large piece of customers fall in detractors segment which means company has to invest more time and effort on these customers. To further classify this detractors segment, company should also consider the number of orders canceled in this segments and what are most common reason for order cancellation. It may include better product description, ease of use, accessibility, correct depiction of product image, and more on the product value that brings to the customer on such purchases. Focusing on canceled invoices to convert into positive order. Converting canceled orders will eventually bring both increased sales and customer loyalty.
Company should also look into order cancellation proportion against the total order placed by customer for each segment of customers.

Furthermore, customer lifetime value can also be calculated to predict the average amount spend by each customer in next 1-2 years. Calculating CLV will also help the marketing team to focus more on such customer who have higher CLV in each segment.



